---
title: "covid 19"
author: "Destiny Calvillo"
date: "7/16/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata$sex <- as.factor(kRdata$sex)
kRdata$result <- as.factor(kRdata$result)
```

### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
dd2 <- kRdata %>% group_by(result,sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

  
```{r}
barchartGC(~result + sex,data=kRdata)
barchartGC(~result + sex,data=kRdata, type="percent")

```

The plots above appear to show that out of the people that ended up dead 60% of them were males and about less than 40% were females. It also that less than 60% of people who are alive were females, and a little more than 40% of people who ended up alive are males. From the percentage bar-chart alone, we can already tell that more men are ending up dead than alive compared to females when looking at the results. 

## Numerical results

```{r}
table1 <- xtabs(~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

When looking at the numerical results, we can see that males are almost as twice as likely to die from Covid-19 compared to females. Out of 100, males make up 64% of covid related deaths compared to only 36% of females. These results correlate to the bar-charts above. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


## Conclusions

